### 2013-11-15 abstract.tex
The abstract template

%%%%%  LaTeX template for submission of abstracts to  MaxEnt 2013         
%%%%%                                                         
%%%%% Use this template to generate a pdf version of your abstract for submission. This is 
%%%%% most easily accomplished using a LaTeX compiler that outputs directly to pdf.
%%%%%                                      
%%%%% Your abstract may be no more that one page in length and that all text and figures must                                                      
%%%%% fit in a 16.0 cm x 23.0 cm frame. 
%%%%%                                          
%%%%% When you submit the pdf version of your abstract on the conference web site
%%%%% you will be ask if your are requesting an Oral presentation or a Poster.  You will                             
%%%%% also be ask it this is a student presentation.                                          
%%%%%

%%%%% Do not change the text size in the line below.
\documentclass[letterpaper,12pt]{article}
%\documentclass[final,onecolumn]{aipproc}
%\layoutstyle{6x9}

%%%%%  If you chose not to include the optional figure then you can delete the line below.
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

%\usepackage{empheq}   %boxes around equations
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}	% places footnotes at page bottom
%\usepackage{epstopdf} 	%essential with TeXShop on Mac
%\usepackage{longtable} 	%long tables
%\usepackage{esint}  	%multiple integral fonts
%\usepackage{stmaryrd} 	%floor, ceiling symbol font

%\hyphenation{MaxEnt}

%%%%% Page size and layout commands. Do not change lines below below.
\textwidth16cm
\textheight23cm
\topmargin-2.0cm
\oddsidemargin0cm
\parindent0pt
%\pagestyle{empty}
%%%%% End of Page size and layout commands.

\begin{document}

%%%%%  The title of your contribution in all capital letters.
\title{TITLE OF YOUR CONTRIBUTION}

%%%%%  Author's name (underline presenter's author)            
\author{\underline{F. Author}$^{1,2}$, S. Author$^{3}$\\
        (1) First Institution\\
        (2) Second Institution\\
        (3) Third Institution\\
        (You might put here additional information: e-mail,
         URL, etc)
       }

\date{date}
\maketitle

%%%%%               Abstract begins here                      
\begin{abstract}
\noindent
In this work we show how to calculate the answer to the Ultimate Question of Life, the Universe and Everything using a combination of Bayes rule and MaxEnt:
\begin{gather}
P(H|D,I) = \frac{P(D|H,I) \, P(H|I)}{P(D|I)}
\\
\mathcal{H} = - \sum_i p_i \ln \frac{p_i}{q_i}
\end{gather}
and a very large organic computer.

%%%%%       OPTIONAL: Figure (delete if not used)                                                 
%%%%%       Do not add a figure caption.
%%%%%       The conference program will not be printed in colour so colour figures should not be used.
%%%%%       The example figure is in pdf format.

%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=70mm]{roo}
%\end{center}
%%\caption{blah blah.}
%%\label{fig:roo}
%\end{figure}
%%%%%        End of Figure

%%%%%       OPTIONAL: References (delete  if not used)   
%%%%%       Separate each reference with a blank line.                                                                
\medskip\noindent{References: }

[1] A. Dent et al. Phys. Rev. {\bf E99}, 99998 (2009).

[2] Z. Beeblebrox et al. Phys. Rev. {\bf E99}, 99999 (2009).
%%%%%        END of References

%%%%%       OPTIONAL: Key Words (delete if not used)   
\medskip\noindent{Key Words: }
{Jaynes, Deep Thought, Marsupials}
%%%%%        END of Key Words

\end{abstract}
\thispagestyle{empty}
\end{document}

\endinput